The human in the loop: volunteer-based metacomputers as a socio-technical system

### 2016-04-24 Alife16/paper.tex
Minor error
ref

This loose connection provides a low-overhead way to connect desktop
experiments with volunteer-based ones, with all of them contributing
to the pool, but every one of them working as separate island carrying
out their own evolutionary algorithms. This is why NodIO is proposed
mainly as a {\em complement} to existing resources such as desktop
systems or laptops. As long as it provides a non-null computational
capability that can help existing resources find the solution faster
it will have found its purpose. The main use case is someone setting
up NodIO in the cloud, writing a fitness evaluation function and
running a client from his or her own computer, but requesting help in
social networks for additional resources. That is why, in this system,
we have considered the whole {\em social} aspect in the design, with
issues related to security, trust and privacy among others. The
computing system becomes a {\em sociotechnical} system}
\citep{vespignani2009predicting}. In this paper, we are going to focus
on measuring the response of users to experiments, that is, the time
they spend running it, but at the same time we will also focus on the
technical aspects of the server and how these might change the
behavior of users, improving the capability of the system.

@@

\begin{figure}[!htb]
\centering
\includegraphics[width=0.32\linewidth]{weibull-puts-openshift-4-4.png}
\includegraphics[width=0.32\linewidth]{weibull-puts-openshift-4-24.png}
\includegraphics[width=0.32\linewidth]{weibull-puts-openshift-7-31.png}
\includegraphics[width=0.32\linewidth]{weibull-fit-cache=128.png}
\includegraphics[width=0.32\linewidth]{weibull-fit-cache=64.png}
\includegraphics[width=0.32\linewidth]{weibull-fit-cache=32.png}
\caption{Number of {\tt PUT}s per unique IP and fit to a Weibull
  distribution (in red). From left to right and top to bottom, experiments
  4/4, 4/24 and 7/31 and new experiments with cache=128, 64, 32.} 
% Plotted with ../data/plot-zipf-openshift.R
\label{fig:puts:os}
\end{figure}
%
A similar exponential distribution also appears if we rank HTTP {\tt
  PUT}s, equivalents to the number of 
generations divided by 100, or to evaluations divided by 12800,
contributed by every user, which is shown 
in Figure \ref{fig:puts:os}. These results show a Zipf-like behavior,
that is, a power law with a small {\em bump} in the lowest
values. After testing the Generalized Extreme Value distribution and
failing for the new batch of experiments, we have fitted it to a
Weibull distribution \cite{thoman1969inferences} with the resulting
parameters shown in Table \ref{tab:puts:os}. 
%
\begin{table}
\caption{Weibull distribution  parameters of the fit of
  the number of {\tt PUT}s per unique IP. \label{tab:puts:os}}
\begin{center}
\begin{tabular}{l|cc}
\hline
Experiment  &  Scale $\sigma$ & Shape $\xi$ \\
\hline
4/4 &  43.07 $\pm$ 5.80 &  0.57 $\pm$ 0.03 \\
4/24 & 22.97 $\pm$ 1.57 & 0.66 $\pm$  0.02  \\
7/31 &  53.18 $\pm$ 7.77 &  0.54 $\pm$ 0.03   \\
\hline
Cache 128 & 205.28 $\pm$ 32.10 & 0.77 $\pm$ 0.07 \\ 
Cache 64 & 178.99 $\pm$ 36.44 & 0.60 $\pm$ 0.05 \\ 
Cache 32 & 168.15 $\pm$ 49.28 & 0.57 $\pm$ 0.07 \\
\hline
% from plot-zipf-openshift.R and ips-puts-alife.R
\end{tabular}
\end{center}
\end{table}
%
The inverse
Weibull distribution is a special case of the GEV distribution
in those papers, and appears usually in natural sciences and
artificial life, usually related to decay. It has been frequently
fitted to volunteer computing frameworks such as SETI@home
\citep{javadi2009mining}. The model that user behavior follows can be
explained straighforwardly: when users visit the page, it draws their
attention for a limited amount of time. They give it a chance for a
few seconds. If something there amuses them or they can engage in a
conversation about it, they stay for a while longer, otherwise, they
leave. The {\em scale} parameter, which is around 20-40 in the first
batch for the 40 Trap problem and between 160 and 200 in the 60 traps
problem, depends mainly on the maximum number of generations people
leave it running. Since it finishes or stalls after a number of
generations shown in Figure \cite{tab:summary:os}, volunteers just
leave after that. Curiously enough, the scale parameter is roughly
twice the median number of {\tt PUT}s per experiment, showing that, on
average, the most loyal users reload the page twice after finishing or
after seeing the evolution does not progress. This rule of thumb
breaks down with the last experiment, however, which is interesting by
itself, too.

>>>>

This loose connection provides a low-overhead way to connect desktop
experiments with volunteer-based ones, with all of them contributing
to the pool, but every one of them working as separate island carrying
out their own evolutionary algorithms. This is why NodIO is proposed
mainly as a {\em complement} to existing resources such as desktop
systems or laptops. As long as it provides a non-null computational
capability that can help existing resources find the solution faster
it will have found its purpose. The main use case is someone setting
up NodIO in the cloud, writing a fitness evaluation function and
running a client from his or her own computer, but requesting help in
social networks for additional resources. That is why, in this system,
we have considered the whole {\em social} aspect in the design, with
issues related to security, trust and privacy among others. The
computing system becomes a {\em sociotechnical system}
\citep{vespignani2009predicting}. In this paper, we are going to focus
on measuring the response of users to experiments, that is, the time
they spend running it, but at the same time we will also focus on the
technical aspects of the server and how these might change the
behavior of users, improving the capability of the system.

@@

\begin{figure}[!htb]
\centering
\includegraphics[width=0.32\linewidth]{weibull-puts-openshift-4-4.png}
\includegraphics[width=0.32\linewidth]{weibull-puts-openshift-4-24.png}
\includegraphics[width=0.32\linewidth]{weibull-puts-openshift-7-31.png}
\includegraphics[width=0.32\linewidth]{weibull-fit-cache=128.png}
\includegraphics[width=0.32\linewidth]{weibull-fit-cache=64.png}
\includegraphics[width=0.32\linewidth]{weibull-fit-cache=32.png}
\caption{Number of {\tt PUT}s per unique IP and fit to a Weibull
  distribution (in red). From left to right and top to bottom, experiments
  4/4, 4/24 and 7/31 and new experiments with cache=128, 64, 32.} 
% Plotted with ../data/plot-zipf-openshift.R
\label{fig:puts:os}
\end{figure}
%
A similar exponential distribution also appears if we rank HTTP {\tt
  PUT}s, equivalents to the number of 
generations divided by 100, or to evaluations divided by 12800,
contributed by every user, which is shown 
in Figure \ref{fig:puts:os}. These results show a Zipf-like behavior,
that is, a power law with a small {\em bump} in the lowest
values. After testing the Generalized Extreme Value distribution and
failing for the new batch of experiments, we have fitted it to a
Weibull distribution \cite{thoman1969inferences} with the resulting
parameters shown in Table \ref{tab:puts:os}. 
%
\begin{table}
\caption{Weibull distribution  parameters of the fit of
  the number of {\tt PUT}s per unique IP. \label{tab:puts:os}}
\begin{center}
\begin{tabular}{l|cc}
\hline
Experiment  &  Scale $\sigma$ & Shape $\xi$ \\
\hline
4/4 &  43.07 $\pm$ 5.80 &  0.57 $\pm$ 0.03 \\
4/24 & 22.97 $\pm$ 1.57 & 0.66 $\pm$  0.02  \\
7/31 &  53.18 $\pm$ 7.77 &  0.54 $\pm$ 0.03   \\
\hline
Cache 128 & 205.28 $\pm$ 32.10 & 0.77 $\pm$ 0.07 \\ 
Cache 64 & 178.99 $\pm$ 36.44 & 0.60 $\pm$ 0.05 \\ 
Cache 32 & 168.15 $\pm$ 49.28 & 0.57 $\pm$ 0.07 \\
\hline
% from plot-zipf-openshift.R and ips-puts-alife.R
\end{tabular}
\end{center}
\end{table}
%
The inverse
Weibull distribution is a special case of the GEV distribution
in those papers, and appears usually in natural sciences and
artificial life, usually related to decay. It has been frequently
fitted to volunteer computing frameworks such as SETI@home
\citep{javadi2009mining}. The model that user behavior follows can be
explained straighforwardly: when users visit the page, it draws their
attention for a limited amount of time. They give it a chance for a
few seconds. If something there amuses them or they can engage in a
conversation about it, they stay for a while longer, otherwise, they
leave. The {\em scale} parameter, which is around 20-40 in the first
batch for the 40 Trap problem and between 160 and 200 in the 60 traps
problem, depends mainly on the maximum number of generations people
leave it running. Since it finishes or stalls after a number of
generations shown in Figure \ref{tab:summary:os}, volunteers just
leave after that. Curiously enough, the scale parameter is roughly
twice the median number of {\tt PUT}s per experiment, showing that, on
average, the most loyal users reload the page twice after finishing or
after seeing the evolution does not progress. This rule of thumb
breaks down with the last experiment, however, which is interesting by
itself, too.
